---
title: "Problem Set 2"
author: "Due 31 January"
output:
  pdf_document: default
  word_document: default
fontsize: 12pt
---

```{r setup, include=FALSE}
  knitr::opts_chunk$set(
    echo = TRUE, 
    eval = TRUE
  )
```
```{r}
library(knitr)
library(ggplot2)
```

Answer the questions below to the best of your ability. Write clearly, and format your tables and visuals appropriately. You must use `R Markdown` to compose and compile your work. For full credit, `echo` all code chunks, and include your `setup` chunk. Submit your work in hard copy at the beginning of class. 

You need the [Global Greenspace Indicator Data](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/TMWYHB) for this assignment. Review the `README.txt` file for information about the data, variables, etc.

1. Show me that you're all set on GitHub. Create a public repository named `PSet2`. Clone it, include all your project files (including your `.Rmd` and `.pdf` files) for the work below, and commit/push your work to your repository. Include the link to your repo as your answer to this question.  

2. The script below doesn't work. Type the corrected code chunk into your problem set. Annotate any line you correct to note your fix (i.e. `# unbalanced parentheses`). *There are more than five errors.*

```{r}
df = readr::read_csv('greenspace_data_share.csv')
```

```{r}
library(tidyverse)

# open my data
 gspace = read_csv("greenspace_data_share.csv") #1 Changed read_csv to read.csv. A valid function
  
# summarize average urban greenspace by region
  table = 
    gspace |>
    group_by(Major_Geo_Region) |> #2 Added pipe function
    summarise(
      obs = n(), #3 Added in a comma 
      avg = mean(annual_avg_2020), #4 Corrected annual_avg_2020 variable name
      weightedavg = mean(annual_weight_avg_2020) #5 Combined weighted and avg text
    )
  
# output as table

knitr::kable(table, format = "markdown")
#6 Corrected data frame from gspace to table
#7 Updated argument from digits to format
#8 Added in knitr 
```

3. How many urban areas does the greenspace data cover?
 
  #There are 1038 urban areas

4. In a couple of sentences and with reference to a well-formatted tabulation, describe the greenspace classification scores for urban areas in 2021.

```{r}
 table2= 
    gspace |>
    group_by(indicator_2021) |> 
    summarise(
      obs = n(), 
      avg = mean(annual_avg_2021), 
      weightedavg = mean(annual_weight_avg_2021) 
    )
  
knitr::kable(table2, format = "markdown")
```
- 6 Classifications of NDVI (vegetation coverage), from Exceptionally Low to Very High.
- 73% of urban areas have low vegetation coverage (Exceptionally low, Very low, and Low).
- Only 6% of urban areas have High or Very High coverage.
- Exceptionally low urban areas have an average of 13% coverage during the year. 
- On the other hand, 42% of High urban areas have coverage. Marking a 29% difference.
- There is only one city with Very High coverage which is Kandy in Sri Lanka. This city is heavily forested in an agricultural region and is a UNESCO protected site, all leading to its exceptional score.


5. Report the number of urban areas that satisfy the conditions below. Either write your code inline or echo the code that generated the answer.   

    a. Scored `High` or above for greenspace in 2015.  

```{r}
  filter(gspace, indicator_2015 == "High" |  indicator_2015 == "Very High") |>
    summarise(
      obs = n())
  
  count(gspace,indicator_2015)
```  
  #There are 66 urban areas  
  
    
    b. Scored `Exceptionally Low` at any point in the years covered.  
  
```{r} 
filter(gspace, indicator_2010 == "Exceptionally Low" |  indicator_2015 == "Exceptionally Low" |             indicator_2020 == "Exceptionally Low" | indicator_2021 == "Exceptionally Low") |>
    summarise(
      obs = n())
```
  #There are 240 urban areas scored as Exceptionally Low from 2010, 2015, 2020, and 2021.
  
    c. Urban areas in arid climate that became greener (as measured by annual weighted average) from 2010 to 2020. 
```{r}  
gspace <- transform(gspace,
  greener = annual_weight_avg_2020 - annual_weight_avg_2010
  )    
  
filter(gspace, Climate_region == "Arid" , greener > 0) |>
    summarise(
      obs = n())
```
  #225 urban areas became greener from 2010 to 2020.

6. How many urban areas became less green (measured by annual average) from 2010 to 2021? Were these changes concentrated in a particular geographic or climate region? Explain (with evidence, of course).

```{r}
gspace <- transform(gspace,
  greener2 = annual_avg_2021 - annual_avg_2010
  )    
  
filter(gspace, greener2 < 0) |>
    summarise(
      obs = n())
```

```{r}
table4= 
    gspace |>
    group_by(Major_Geo_Region) |> 
    filter(greener2 < 0) |>
    summarise(
      obs = n())
```
```{r}
table5= 
    gspace |>
    group_by(Climate_region) |> 
    filter(greener2 < 0) |>
    summarise(
      obs = n())
```
      
#6A 128 urban areas became less green from 2010 to 2021
#6B The changes are in Europe and termperate regions as seen in the tables. 

7. Present a histogram showing the change in greenspace (annual average) from 2010 to 2021. Note that you will need to create a new variable equal to this difference.  
```{r}
hist(gspace$greener2, 
     main = "Change in greener2 from 2010 to 2021",
     xlab = "Change in greener2",
     ylab = "Frequency",
     col = "red",border = 'green')
```

8. Present a scatter plot of population weighted greenspace in 2021 over the greenspace in 2010. 
```{r}
ggplot(df, aes(x = annual_weight_avg_2010, y = annual_weight_avg_2021)) +
geom_point () +
labs(
    x = 'Annual Weight Average 2010',
    y = 'Annual Weight Average 2021'
  )  
```

```{r}
df$slope <- (df$annual_weight_avg_2021 - df$annual_weight_avg_2010) / df$annual_weight_avg_2010

ggplot(df, aes(
      x = annual_weight_avg_2010,
      y = annual_weight_avg_2021)
      ) + 
  geom_point(aes
              (color = ifelse(slope > 0,'Moregreenspace',
                ifelse(slope < 0, 'Lessgreenspace', 'same')))) +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted", color = "pink") +
  labs(title = 'Scatter Plot', 
    x = 'Annual Weighted Average 2010',
    y = 'Annual Weighted Average 2021')+
  scale_color_manual(values = c('same' = "blue",'Lessgreenspace'= "green",
                      'Moregreenspace'='red'), name = 'Legend')
```
                      
**BONUS OPPORTUNITY**: Use color-coding to differentiate urban areas that added versus lost greenspace in that time. Then include a 45-degree line to further highlight the change.
